---
title: "Modding Rome Total War (Avatar, The Last Airbender mod) using R"
always_allow_html: yes
output: 
  md_document:
    variant: markdown_github
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This blog post will be on a very weird mash-up of a few of my interests, Rome: Total War, Avatar: The Last Airbender, and of course the R programming language! I'll introduce these drastically different interests first...

## __What is Rome: Total War?__ 

- A turn-based strategy RPG game set in the early years of the Roman Republic that was released back in 2004. There's both a campaign map where you do management (akin to say Civ or EU games) but a major difference to those games is that you can also fight the battles yourself on the battle map with the troops you've recruited. 

![]()

- It's probably the game I've played the most, ever. Yes, even more than FIFA actually (I haven't played FIFA in nearly a decade now actually...). I have the original CD version that I bought when I lived in Israel, the Steam version which I play on today, and I also got the Re-mastered version as that's the version that works with the Avatar mod.

## __What is Avatar: The Last Airbender?__

- An animated TV series from the mid-2000s about a group of teenagers from various "tribes/nations" who set out to restore balance and harmony to their world which as been ravaged by war for the past 100 years.

- It's actually better if you just watch the 45-second opening as it sets the scene extremely well in a succinct manner:

[[avatar-intro](https://flxt.tmsimg.com/assets/p186109_b_v8_aa.jpg)](https://www.youtube.com/watch?v=IAdd8RSg86g)

- It got pretty popular (again) during the pandemic I feel... Most likely because the people who watched it when they were kids are right around the late 20s/early 30s-ish age that are pretty active on social media and people wanted some comfort-shows to watch in those uncertain times.

## __What is the Avatar: The Last Airbender mod for Rome: Total War?__

I was working on this during fall of last year but then I got busy (J.League review, World Cup, etc.) so I never wrote up the blog post in full. A lot of the actual modding is done outside of R, but as always I was able to find ways to use R to my advantage to improve this mod.

"and now for something completely different..."

Let's get started!

## Web-scraping Avatar characters to add into the game.

When I first acme across this mod, I realized pretty quickly that there weren't a whole lot of character names for each of the tribes. So after like 50 turns or so of playing the game, you wind up having a dozen or so "Katara"s running around in your Water Tribe kingdom or 50 "Zuko"s in your Fire Nation Empire. Therefore I took it upon myself to add more character names by extracting names and gender from the [Avatar fandom wiki](https://avatar.fandom.com/wiki/Avatar_Wiki).

First, let's load some R packages: 

```{r}
library(rvest)   ## web-scraping
library(polite)  ## web-scraping, politely
library(dplyr)   ## clean data
library(purrr)   ## clean data and iteration
library(stringr) ## string/text cleaning
```

Before I jumped into creating a big workflow, I wanted to test how to grab the data I need based on one page.

```{r}
## Check out robots.txt and see if URL link is scrape-able
session2 <- bow("https://avatar.fandom.com/wiki/Azula")

## Find the character info box on the right corner and 
## grab all the text, then extract the gender identifying words
gender_is <- scrape(session2) %>%
  html_nodes(".portable-infobox") %>%
  html_text() %>%
  ## Some characters are non-binary or gender isn't listed
  stringr::str_extract("Female|Male|Man|Woman") 

gender_is
```

Great! After flipping through a few character pages at random, this is how these pages are set up so I should be able to apply the above code to **every** character page to extract the character name and their gender. 

NOTE: Since the time I actually worked on this last October, there's been quite a few weird changes in the pages where instead of simply listing gender as "Male" or "Female" or otherwise, some pages keep the "Male"/"Female" set while others have "Man" or "Woman" instead... so the consistency across character pages got worse, wonderful!

So to expand this search across all characters, I tried to find a pages that listed them all. There wasn't one but there are a couple of pages that lists all characters split by tribe/nation which is still more than enough for what I wanted to do.

```{r}
earth_url <- "https://avatar.fandom.com/wiki/Category:Earth_Kingdom_characters"
fire_url <- "https://avatar.fandom.com/wiki/Category:Fire_Nation_characters"
water_url <- "https://avatar.fandom.com/wiki/Category:Water_Tribe_characters"
air_url <- "https://avatar.fandom.com/wiki/Category:Air_Nomad_characters"
```

So from here, I need to grab:

- The character's name
- The URL link to that character's page

```{r, eval=FALSE}
## one nation
bender_url <- earth_url
bender_label <- str_extract(bender_url, "Fire|Earth|Water|Air")

cat(paste0("\nStarting: ", bender_label, "!\n"))
## Get name and page link
session <- bow(bender_url)

char_link <- scrape(session) %>%
  html_nodes(".category-page__member-link") %>%
  html_attr("href")

char_name <- scrape(session) %>%
  html_nodes(".category-page__member-link") %>%
  html_text()

base_url <- "https://avatar.fandom.com/"

earth_df <- data.frame(
  char_link,
  char_name
) %>%
  ## Strip out 'Category:' pages that aren't single character page links
  filter(!str_detect(char_link, "Category")) %>%
  mutate(char_link = paste0(base_url, char_link))

cat("\nNames and page link done!\n")
```

```{r, echo=FALSE}

```

```{r}
earth_df
```

Then use that information to:

- Jump into each character page to find their gender
- Bind all individual character info into one data.frame and save that as an `.RDS` file

```{r, eval=FALSE}
## find all genders at once
earth_namegender_df <- map(
  earth_df$earth_link,
  ~ get_gender(data = earth_df, link = .x)
) %>%
  purrr::reduce(earth_namegender_df, bind_rows)

saveRDS(earth_namegender_df, file = "data/earth_namegender_df.RDS")
```

Of course, after running through one nation's worth of character pages I can then standardize it as one big function to take any of the character list URLs.

```{r}
get_character_info <- function(bender_url) {
  bender_label <- str_extract(bender_url, "Fire|Earth|Water|Air")

  cat(paste0("\nStarting: ", bender_label, "!\n"))
  ## Get name and page link
  session <- bow(bender_url)

  char_link <- scrape(session) %>%
    html_nodes(".category-page__member-link") %>%
    html_attr("href")

  char_name <- scrape(session) %>%
    html_nodes(".category-page__member-link") %>%
    html_text()

  base_url <- "https://avatar.fandom.com/"

  char_df <- data.frame(
    char_link,
    char_name
  ) %>%
    ## Strip out 'Category:' pages that aren't single character page links
    filter(!str_detect(char_link, "Category")) %>%
    mutate(char_link = paste0(base_url, char_link))

  cat("\nNames and page link done!\n")

  ## Get gender
  get_gender <- function(data, link) {
    cat(paste0("\nStarting: ", link))
    session2 <- bow(link)

    gender_is <- scrape(session2) %>%
      html_nodes(".portable-infobox") %>%
      html_text() %>%
      stringr::str_extract("Female|Male|Man|Woman") ## Some are "Non-binary" or gender isn't listed

    cat(paste0("\nGender: ", gender_is))

    final_df <- data %>%
      filter(char_link == link)

    if (class(gender_is) != "character") {
      final_df <- final_df %>%
        mutate(gender = NA_character_)
    } else {
      final_df <- final_df %>%
        mutate(gender = gender_is)
    }

    cat("\nDone!\n")

    return(final_df)
  }

  char_namegender_df <- map(
    char_df$char_link,
    ~ get_gender(data = char_df, link = .x)
  ) %>%
    purrr::reduce(bind_rows)

  saveRDS(char_namegender_df, file = here::here(paste0("data/", bender_label, "_namegender_df.RDS")))

  cat(paste0("\nScript for: ", bender_label, " done!\n"))
}
```

And from here I basically shoved all these character names by nation/tribe into the game's character files using NotePad++ since most of the game is built on simple text files! It's very easy to mod this part of the game, all of the model/map rendering and stuff is obviously much harder though.

## Generating a "turns-per-year" script through R's text generation capabilities

The way the campaign is set up in this game is that the seasons alternate every two turns (summer, winter) and two turns make up a full year. However, this can create problems where your favorite characters age really quickly and you don't get to spend a whole lot of time developing their skills and role-playing as them before they kick the bucket. So throughout all the different total war games, there have been mods or scripts made to extend the "turns-per-year" mechanic in the campaign (4, 8, 16, turns-per-year, etc. depending on the setting/size of map/etc.).

This is especially true for a mod like Avatar: The Last Airbender. While, all the protagonists start off as young teenagers, the fact that they can die in only 100 or so turns makes the campaign a bit drab once they all bite the dust and are replaced by a bunch of randomly generated characters. 




First I set up the base data.frame. `cons_date` is the year repeated 4 times each to represent 4-turns-per-year. Most people won't play up to 200 turns but just in case I extended it to 800 turns in the `turn_num` variable. Keep in mind that we're starting from turn `0`/zero and not turn `1`.

I create a separate vector that makes sure to change the season every 4 turns starting from turn 2. Then I set a separate vector representing which turns will be the "winter" turns, starting from turn 3 (since we're first changing the season after turn 2). 

```{r}
## Generate turn numbers and other vars
turn_df <- data.frame(
  cons_date = rep(1:200, each = 4),
  turn_num = seq(0, 799)
)

season_change_turn <- seq(2, 800, by = 4)
winter_turn <- seq(3, 800, by = 4)
```

From here I open up a file to write my code into: `4tpy_AV.txt`. From there I use `cat()` calls to push text into the file that I created, first some comments describing what this script is doing.

```{r}
fileName <- sprintf("4tpy_AV.txt")

message(sprintf("Writing %s...", fileName))
r <- file(fileName, open = "w")

cat(file = r, "; GENERATED by RN-7 via R script\n\n")

cat(file = r, "script\n\n")

cat(file = r,
    "; 4 turns per year script\n")

cat(file = r,
    "; starting year: 1, ending year: 200\n\n")

cat(file = r,
    "while I_TurnNumber < 800\n\n")

cat(file = r,
    "
    suspend_during_battle on\n")
```

From here, it's a giant loop of setting the text snippet for each turn/season in a `for` loop.

```{r}
## for loop
for (i in 1:nrow(turn_df)) { # 1:nrow(turn_df)
  # print(turn_df[i, 1]) # cons_date
  # print(turn_df[i, 2]) # turn_num

  ## check if 'i' is season changing turn
  if (turn_df[i, 2] %in% season_change_turn) {
    cat(file = r,
        sprintf("
    console_command date %s
    console_command season summer
    monitor_event FactionTurnEnd FactionType slave
        if I_TurnNumber = %s
                console_command season summer
        end_if
    terminate_monitor
    end_monitor
    while I_TurnNumber = %s
    end_while\n\n",
    turn_df[i, 1], turn_df[i, 2], turn_df[i, 2]
        )
    )
  } else if (turn_df[i, 2] %in% winter_turn) {
    cat(file = r,
        sprintf("
    console_command date %s
    console_command season winter
    while I_TurnNumber = %s
    end_while\n\n",
    turn_df[i, 1], turn_df[i, 2]
        )
    )
  } else {
    cat(file = r,
        sprintf("
    console_command date %s
    console_command season summer
    while I_TurnNumber = %s
    end_while\n\n",
    turn_df[i, 1], turn_df[i, 2]
        )
    )
  }
}

## Close connection to file ----
close(r)
```

So I run the entire script and boom! 


Unfortunately, I don't know enough about modding to make the UI track the "actual" date rather than just the year. So in the campaign map you just get:

- "Year 1 (Summer icon)", "Year 1 (Summer icon)", "Year 1 (Summer icon)", "Year 1 (Winter icon)", 
- "Year 2 (Summer icon)", "Year 2 (Summer icon)", "Year 2 (Summer icon)", "Year 2 (Winter icon)", 
- "Year 3 (Summer icon)" ... and so on...




This is how set up my soccer data analysis workflow. Each year or half-a-season I am creating the same visualizations from similar data sets. The only thing that changes is the name of the folders or the name of the data sets (ex. `2022_data` to `2023_data`), names of the teams involved (which I can just grab from the incoming data set), etc... so I really only need to define these as variables at the top of my script then carefully make sure that the text I'm generating will create the exact set of data-cleaning/data-visualization code that I need with the new data set.

I can run the script to generate the R script file using the new season//half-a-season//whatever variables, then run the script itself and all my data viz pop up! It saves me a lot of time as I don't have to copy-paste my previous script and then find-replace stuff like `2022` to `2023` and whatnot.





